---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bark: Bayesian Additive Regression Kernels

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/bark)](https://CRAN.R-project.org/package=bark)
[![R-CMD-check](https://github.com/merliseclyde/bark/workflows/R-CMD-check/badge.svg)](https://github.com/merliseclyde/bark/actions)
[![codecov](https://codecov.io/gh/merliseclyde/bark/graph/badge.svg?token=iPCcWEu34R)](https://app.codecov.io/gh/merliseclyde/bark)
[![OpenSSF Best Practices](https://bestpractices.coreinfrastructure.org/projects/7096/badge)](https://bestpractices.coreinfrastructure.org/projects/7096)

<!-- badges: end -->

The bark package implements estimation for a Bayesian nonparametric regression model represented as a sum of multivariate Gaussian kernels as a flexible model to capture nonlinearities, interactions and feature selection.

## Installation

You can install the released version of bark [![](https://www.r-pkg.org/badges/version/bark)](https://cran.r-project.org/package=bark) from [CRAN](https://CRAN.R-project.org) with:

```{r eval=FALSE}
install.packages("bark")
```

And the development version from [GitHub](https://github.com/merliseclyde/bark) with:

```{r eval=FALSE}
require("devtools")
devtools::install_github("merliseclyde/bark")
```
(verify that the branch has a passing  R CMD check badge above)

## Example



```{r example}
library(bark)
set.seed(42)
traindata <- sim_Friedman2(200, sd=125)
testdata <- sim_Friedman2(1000, sd=0)
fit.bark.d <- bark(y ~ .,  
                   data=data.frame(traindata), 
                   testdata = data.frame(testdata),
                   classification=FALSE, 
                   selection = TRUE,
                   common_lambdas = FALSE,
                   printevery = 10^10)

mean((fit.bark.d$yhat.test.mean-testdata$y)^2)
```

bark is similar to SVM, however it allows different kernel smoothing parameters for every dimension of the inputs  $x$ as well as selection of inputs by allowing the kernel
smoothing parameters to be zero.

The plot below shows posterior draws of the $\lambda$ for the simulated data.

```{r}
boxplot(as.data.frame(fit.bark.d$theta.lambda))
```

The posterior distribution for $\lambda_1$ and $\lambda_4$ are concentrated near zero, which leads to $x_1$ and $x_2$ dropping from the mean function.

## Roadmap for Future Enhancements

Over the next year the following enhancements are planned:

* port more of the R code to C/C++ for improvements in speed

* add S3 methods for `predict`, `summary`, `plot` 

* add additional kernels and LARK methods from AOS (2011) paper

* better hyperparameter specification

If there are features you would like to see added, please feel free to create an
[issue in GitHub](https://github.com/merliseclyde/bark/issues)  and we can
discuss!
