---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bark: Bayesian Additive Regression Kernels

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/merliseclyde/bark.svg?branch=master)](https://travis-ci.org/merliseclyde/bark)
[![Codecov test coverage](https://codecov.io/gh/merliseclyde/bark/branch/master/graph/badge.svg)](https://codecov.io/gh/merliseclyde/bark?branch=master)
<!-- badges: end -->

The bark parkage implements estimation for a Bayesian nonparametric regression model represented as a sum of multivariate Gaussian kernels as a flexible model to capture nonlinearities, interactions and feature selection.

## Installation

You can install the released version of bark from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("bark")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("merliseclyde/bark")
```
## Example



```{r example}
library(bark)
traindata <- sim_Friedman2(200, sd=125)
testdata <- sim_Friedman2(1000, sd=0)
fit.bark.d <- bark(traindata$x, 
                   traindata$y, 
                   testdata$x, 
                   classification=FALSE, 
                   type="sd",
                   printevery = 10^10)

mean((fit.bark.d$yhat.test.mean-testdata$y)^2)
```

bark is similar to SVM, however it allows different kernel smoothing parameters for every dimension of the inputs  $x$ as well as selection of inputs by allowing the kernel
smoothing parameters to be zero.

The plot below shows posterior draws of the $\lambda$ for the simulated data.

```{r}
boxplot(as.data.frame(fit.bark.d$theta.lambda))
```

The posterior distibution for $\lambda_1$ and $\lambda_4$ are concentrated at zero, which leads to $x_1$ and $x_2$ dropping from the mean function.
