---
title: "Nonparametric Regression with Bayesian Additive Regression Kernels"
output: rmarkdown::html_vignette
author: "Merlise A Clyde"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Nonparametric Regression with Bayesian Additive Regression Kernels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```





Bayesian Additive Regression Kernel (BARK) models are a flexible Bayesian nonparametric model for regression and classification problems where the unknown mean function is represented as a weighted sum of multivariate Gaussian kernel functions,
\begin{equation}
f(\mathbf{x}) = \sum_j \beta_j  \prod_d \exp( \lambda_d (x_d - \chi_{jd})^2)
\end{equation}
that captures nonlinearities, interactions and feature selection 
using Levy random fields to construct a prior on the unknown function.   Each kernel is centered at location parameters $\chi_{jd}$ with precision parameters $\lambda_d$ - these precision parameters apture the importance of each of the $d$ dimensional predictor variables and by setting a $\lambda_d$ to zero may remove important varialbes that are not important.


## Installation

To get the latest version  of {r bark}, install from github  (needs compilation))

```{r, eval=FALSE}
devtools::install_github("merliseclyde/bark")
```

```{r}
library(bark)
```

## Example

We will illustrate feature selection in a simple simulated example from Friedman

```{r Friedman2}
set.seed(42)
traindata <- sim_Friedman2(200, sd=125)
X.train = scale(traindata$x)
testdata <- sim_Friedman2(1000, sd=0)
```

```{r example-all}
fit.bark.d <- bark(X.train, 
                   traindata$y, 
                   scale(testdata$x, center = attr(X.train,"scaled:center"),
                         scale = attr(X.train,"scaled:scale")),
                   classification=FALSE, 
                   type="d",
                   printevery = 10^10)

mean((fit.bark.d$yhat.test.mean-testdata$y)^2)

```

```{r example-selection}
fit.bark.sd <- bark(X.train, 
                   traindata$y, 
                   scale(testdata$x, center = attr(X.train,"scaled:center"),
                         scale = attr(X.train,"scaled:scale")),
                   classification=FALSE, 
                   type="sd",
                   printevery = 10^10)

mean((fit.bark.sd$yhat.test.mean-testdata$y)^2)

```

bark is similar to SVM, however it allows different kernel smoothing parameters for every dimension of the inputs  $x$ as well as selection of inputs by allowing the kernel
smoothing parameters to be zero.

The plot below shows posterior draws of the $\lambda$ for the simulated data.

```{r}
boxplot(as.data.frame(fit.bark.d$theta.lambda))
```
```{r}
boxplot(as.data.frame(fit.bark.sd$theta.lambda))
```

The posterior distribution for $\lambda_1$ and $\lambda_4$ are concentrated at zero, which leads to $x_1$ and $x_2$ dropping from the mean function.


## Comparison
We will compare {r bark} to two other popular methods, {r BART} and {r SVM} that provide flexible models that are also non-linear in the input variables. 

```{r}
bart.available =  require(BART)
svm.available  =  require(e1071)
```


### Generate data



```{r}
set.seed(42)
n = 500
circle2 = as.data.frame(bark::sim_circle(n, dim = 2))
train = sample(1:n, size = floor(n/2), rep=FALSE)
```

```{r, fig.width=4, fig.height=4 }
plot(x.1 ~ x.2, data=circle2, col=y+1)
```


```{r bark}

circle2.bark = bark(as.matrix(circle2[train, 1:2]), circle2[train, 3], 
                              x.test = as.matrix(circle2[-train, 1:2]),
                    classification = TRUE,
                    type="se")
```

```{r}
#Classify
#
mean((circle2.bark$yhat.test.mean > 0) != circle2[-train, "y"])
```

## SVM

```{r svm}
if (svm.available) {
  circle2.svm = svm(y ~ x.1 + x.2, data=circle2[train,], type="C")
  pred.svm = predict(circle2.svm, circle2[-train,])
  mean(pred.svm != circle2[-train, "y"])
}

```


```{r bart}
if (bart.available) {
  circle.bart = pbart(x.train = circle2[train, 1:2], 
                            y.train =  circle2[train, "y"])
  pred.bart =   predict(circle.bart, circle2[-train, 1:2])
  mean((pred.bart$prob.test.mean > .5) != circle2[-train, "y"])
} 
```


Compare classification across methods.
```{r plots}

plot(x.1 ~ x.2, data=circle2[-train,], pch = y+15, 
     col=(1 + (circle2.bark$yhat.test.mean > 0)),
     main="bark")

if (bart.available) {
  plot(x.1 ~ x.2, data=circle2[-train,], pch = y+15, 
       col= ( 1 + (pred.bart$prob.test.mean > .5)), 
       main="BART")
}

if (svm.available) {
plot(x.1 ~ x.2, data=circle2[-train,], pch = y+15, 
     col= as.numeric(pred.svm), 
     main="svm")
}


```
